---
title: "LLM-Assisted Analysis of Open-Ended Text"
subtitle: "Consumer Car Reviews from Edmunds.com"
format: html
editor: source
---

<!--
to set this up, we draw on a combination of:
https://quarto.org/docs/publishing/github-pages.html
https://quarto.org/docs/projects/quarto-projects.html
https://www.youtube.com/watch?v=wZ9pZOFSCko&ab_channel=JoshuaFrench
-->

<!-- chunk options
#| fig-cap: "Air Quality"

eval	Evaluate the code chunk (if false, just echos the code into the output).
echo	Include the source code in output
output	Include the results of executing the code in the output (true, false, or asis to indicate that the output is raw markdown and should not have any of Quarto’s standard enclosing markdown).
warning	Include warnings in the output.
error	Include errors in the output (note that this implies that errors executing code will not halt processing of the document).
include	Catch all for preventing any output (code or results) from being included (e.g. include: false suppresses all output from the code block).

-->

## Intro

Companies commonly collect unstructured textual data containing insights into consumer preferences and behavior (e.g., customer reviews, feedback surveys, support tickets, or other product feedback channels). Historically, quantifying and analyzing these data has been challenging, in part because traditional topic modeling or natural language processing methods are complex, opaque, and not always insightful. The advent of LLMs, however, has dramatically simplified the process of analyzing unstructured text and improved the quality of the resulting output.

This page provides a demonstration of the process by which LLMs can be used to code and quantify unstructured textual data. The process involves four steps:

1. Define the codes (variables) to be applied to the unstructured textual data.[^note1]
2. Using your preferred LLM's API, submit the unstructured textual data to the LLM alongside prompts for each code. Specify that results be returned in JSON format.
3. Validate the accuracy of the resulting LLM-coded output against a small subset of human-coded observations. If necessary, refine the prompts and/or codes and repeat steps 1-3 until an acceptable accuracy level is reached.  
4. Analyze the resulting quantitative data as desired. 

[tbd: brief discussion of extensions and uses cases]


[^note1]: Another option is to prompt the LLM to identify major themes or topics in sub-sample of unstructured text responses. The analyst can then use these topics as the basis for the codes that will be applied to the full corpus.     


```{r setup}
#| include: false 

library(tidyverse)
library(haven)
library(readr)
library(RColorBrewer)
library(ggplot2)
library(dplyr)
library(readxl)
library(hrbrthemes)
library(lubridate)

```



```{r insheet}
#| include: false 
#| eval: false

# American car brands: ford, stellantis (chrysler, dodge, jeep, and ram), GM (Buick, Cadillac, Chevrolet, and GMC)
# Japanese car brands: toyota, honda, mazda, nissan, mitsubishi, suzuki, subaru, infiniti, lexus, acura
# European car brands: bmw, mercedes benz, volkswagen, audi, volvo, 
# south korean car brands: kia, hyundai


# Define the path to the subfolder containing the .csv files
subfolder_path <- "data"

# List all .csv files in the subfolder
csv_files <- list.files(path = subfolder_path, pattern = paste0("Scra*",".*\\.csv$"), full.names = TRUE)


# Initialize an empty list to store the data frames
data_list <- list()

# Use lapply to read all .csv files and store them in a list
# data_list <- lapply(csv_files, read_csv)

# Loop through each file and read it into a data frame
for (file in csv_files) {
  
  # extract the make from filename
  # Using gsub to cut "Scraped_Car_Review_" and ".csv"
  result <- gsub("data/Scrapped_Car_Reviews_", "", file)  
  result <- gsub("data/Scrapped_Car_Review_", "", result)  
  result <- gsub("data/Scraped_Car_Review_", "", result)  
  result <- gsub("data/Scraped_Car_Review_", "", result)  
  result <- gsub(".csv", "", result)  

  # Read the .csv file & add variable for make
  data <- read_csv(file, show_col_types = FALSE) |> mutate(make = result)
  
  # Store the data frame in the list
  data_list[[file]] <- data
}

df <-bind_rows(data_list)
glimpse(df)


# remove leading words & spaces from date variable

df <- df |>
  mutate(Review_Date = str_replace_all(Review_Date, "on ", ""),
         Review_Date = str_replace_all(Review_Date, "PDT", ""),
         Review_Date = str_replace_all(Review_Date, " PST", ""),
         Review_Date = str_replace_all(Review_Date, "[()]", ""),
         Review_Date = str_trim(Review_Date),
         Review_Date = mdy_hm(Review_Date))

glimpse(df)

# for some reason, mdy_hm didn't convert above

df <- df |>
  mutate(Review_Date = mdy_hm(Review_Date))
glimpse(df)

# add variable for country and a unique ID

df <- df |>
  mutate(index = 1:nrow(df)) |>
  select(-c("...1")) |>
  mutate(country = case_when(
    make %in% c("toyota", "honda", "mazda","nissan","mitsubishi","suzuki","subaru","infiniti","lexus","acura") ~ "Japanese",
    make %in% c("ford", "chrysler", "dodge", "jeep", "ram", "buick", "cadillac", "chevrolet", "gmc", "lincoln", "mercury") ~ "American",
    make %in% c("bmw", "mercedes-benz","volkswagen","audi", "volvo", "porsche", "mini") ~ "European",
    make %in% c("kia", "hyundai") ~ "South Korean",
    TRUE ~ make  # Default value for any unmatched cases
  ))

glimpse(df)
length(unique(df$index))
table(df$country)


df <- df |>
  select(-Author_Name)


save(df, file = "data\\car_reviews.RData")


```


## Data

We'll use data on car reviews scrapped from [Edmunds.com](www.edmunds.com) and made publicly available on [Kaggle](www.kaggle.com). For each review, the dataset provides the date, vehicle brand, review text and title, and the customer's overall satisfaction rating on a scale of 0 to 5 for. The data cover approximately 206,000 reviews submitted between 2004 and 2018. The goal is to code the sentiment of each review (setting aside for time being thatalready have the reviewer's numeric satisfaction rating), as well as various metrics that we might be interested in tracking overtime, such as whether the reviewer references the vehicle's reliability, safety, drive characteristics, or other variables of interest.  


```{r}

# load packages
library(tidyverse)
library(dplyr)
library(ggplot2)

# load & view data
load("data\\car_reviews.Rdata")
glimpse(df)


```



### Descriptives

The following plot summarizes the average rating (on a scale of 0 to 5) overtime, disaggregated by automaker region:

```{r}
#| warning: false
#| results: 'hide'   # Suppresses printed output (but shows plots)


df.bymonth <- df |>
  mutate(reviews.month = round_date(Review_Date, "month")) |> # reviews by month-year
  #mutate(reviews.month = round_date(Review_Date, "3 months")) |> # reviews by quarter
  group_by(country, reviews.month) |>
  summarise(
    mean.rating = mean(Rating, na.rm=TRUE),
    num.reviews = n()
  )

p <- ggplot(df.bymonth, aes(x=reviews.month, y=mean.rating, group=country, color=country)) +
  geom_smooth(method = "loess", span = .25) +
  theme_minimal()

p


```


## LLM-Assisted Coding

First, let's familiarize ourself with the type of data we'll be classifying. Here's a random sample of five reviews: 

```{r}
#| echo:  FALSE
#| warning: FALSE

library(kableExtra)

df |>
  filter(nchar(Review) < 300 & nchar(Review) > 150) |> # for readability, filter out very long reviews 
  slice_sample(n = 5) |> 
  select(make, Review_Title, Review) |>
  #knitr::kable(, "html")
  knitr::kable(format = "html", escape = FALSE) |>  # escape=FALSE allows HTML in cells
  kable_styling(bootstrap_options = c("striped", "hover")) |>
  scroll_box(width = "100%", height = "300px")  # Adds scroll if needed

```


<!--
We'll use the @ellmer package to code the data. 
-->

\vspace{1em}

We'll use the [**ellmer**](https://ellmer.tidyverse.org/) package to code the data. ellmer supports a wide variety of LLM providers; for now, we'll use OpenAI's GPT-4o model. 

<!--
Note: for first time, need to add your open_api_ai key as an environmental variable. To edit your environment, open Renvirion via: usethis::edit_r_environ()
-->


```{r}
#| warning: false

library(ellmer)

chat <- chat_openai(
  model = "gpt-4o-mini",
  system_prompt = "You are a friendly but terse assistant.",
)

# confirm chat is connected to API
chat$chat("Is R a functional programming language?")

```


Define the prompts that will be used to code the data:

```{r}
prompt<-paste("Please car review the an owner submitted to Edmunds.com, a popular website for reviewing cars. Review the response carefully and code the following variables:")

response_coded <- type_object(
  prompt,
  sentiment = type_enum("Extract the sentiment of the review", c("positive", "neutral", "negative")),
  gpt_Rating = type_number("Based on this review, how do you think the owner would rate their vehicle? Use a numeric scale from 0 to 5, where 0 is very negative, and 5 is very positive."),
  reliability = type_integer("1 if the owner references the vehicle's reliability in their review, 0 otherwise"),
  utility = type_integer("1 if the owner references the vehicle's utility features in their review, 0 otherwise"),
  driving = type_integer("1 if the owner references the vehicle's driving characteristics in their review, 0 otherwise"),
  comfort = type_integer("1 if the owner references the vehicle's comfort in their review, 0 otherwise"),
  interior = type_integer("1 if the owner references the vehicle's interior features in their review, 0 otherwise"),
  fuel_efficiency= type_integer("1 if the owner references the vehicle's fuel efficiency in their review, 0 otherwise"),
  technology = type_integer("1 if the owner references the vehicle's technology features in their review, 0 otherwise")
)


# Demo structured-data function on a single review

text <- df |>
  slice_sample(n = 1) |> 
  select(Review) |>
  as.character()

text

chat <- chat_openai()
coded.review <- chat$extract_data(text, type = response_coded)
str(coded.review)



```


The next step would be to loop through the entire corpus, coding each observation one-by-one. However, with over $206,000$ observations and about 2 seconds for each query to the API, this code would take about 5 days to run. For purposes of this demonstration, we'll use a sample of $7,000$ randomly sampled observations: 


```{r}
#| eval: false
# code a subset of responses
df.sub <- df |>
  slice_sample(n = 7000)

result = NULL
for (i in 1:nrow(df.sub)) {
  out <- chat$extract_data(paste(df.sub[i,"Review"]), 
                          type = response_coded)
  result <- rbind(result,
                  cbind(df.sub[i,"index"],as.data.frame(out)))
}
glimpse(result) 

```


```{r}
#| eval: false
#| include: false

# code to run just once, to code the data ... 


# randomly assign groups

categories <- 1:30

# Randomly assign categories to each observation
set.seed(123)
df <- df %>%
  mutate(
    # Assign random category with equal probability
    sample_group = sample(categories, size = n(), replace = TRUE)
    )



df.sub <- df |>
  filter(sample_group == 2)

# re-initiate connection to API
chat <- chat_openai(
  model = "gpt-4o-mini",
  system_prompt = "You are a friendly but terse assistant.",
)

# confirm chat is connected to API
chat$chat("Is R a functional programming language?")

#to do: reinitiate every 100 requests


result = NULL
for (i in 1:1000) {
  out <- chat$extract_data(as.character(df.sub[i,"Review"]), 
                          type = response_coded)
  result <- rbind(result,
                  cbind(df.sub[i,"index"],as.data.frame(out)))
  print(i) # iteration counter
  #Sys.sleep(2)
  
  if(i %% 5 == 0) Sys.sleep(2) # Pause every 5 requests
  if(i %% 100 == 0) {  # re-initiate connection to API every 100 obs
  chat <- chat_openai(
    model = "gpt-4o-mini",
    system_prompt = "You are a friendly but terse assistant.",
    )
}
}

glimpse(result)

write_csv(result, file = "data\\car_reviews_coded_sample2_1.csv")



# compile all coded batches and save as .csv

  file_list <- list.files(path = "data", 
                       pattern = "coded_sample\\d+",
                       full.names = TRUE) 
 
  # Initialize an empty list to store data
  data_list <- list()
  
  # Read each file and add to the list
  for (i in seq_along(file_list)) {
    message("Reading file ", i, " of ", length(file_list), ": ", basename(file_list[i]))
    
    # Read the CSV file and add filename as a column
    data_list[[i]] <- read_csv(file_list[i], show_col_types = FALSE) #%>%
      #mutate(source_file = basename(file_list[i]))
  }
  
  # Combine all data frames by rows
  combined_data <- bind_rows(data_list)
  dim(combined_data)
  
write_csv(combined_data, "data\\car_reviews_coded_sample.csv")  


```


Having run the code, we now have a fully-coded sample of 7000 reviews:

```{r}
#| echo: false
#| warning: false

df.coded.sample <- read_csv("data\\car_reviews_coded_sample.csv",
                     show_col_types = FALSE) # suppress column type messages
glimpse(df.coded.sample)

# keep only matching rows in coded sample and full sample
df.coded <- inner_join(df.coded.sample, df, by = "index")
#glimpse(df.coded)

```

Which we can then visualize as follows:

```{r}
#| warning: false

# visualize the data

df.bymonth <- df.coded |>
  mutate(reviews.month = round_date(Review_Date, "month")) |> # reviews by month-year
  group_by(reviews.month) |>
  summarise(
    mean.reliability = mean(reliability, na.rm=TRUE),
    mean.driving_characteristics = mean(driving, na.rm=TRUE),
    mean.comfort = mean(comfort, na.rm=TRUE),
    mean.fuel_efficiency = mean(fuel_efficiency, na.rm=TRUE),
    mean.technology = mean(technology, na.rm=TRUE)
  ) |> 
  pivot_longer(
  cols = starts_with("mean."),  # columns to reshape
  names_to = "car_characteristic",          # new column name for the variable names
  names_prefix = "mean.",       # remove prefix from column names
  values_to = "proportion"            # new column name for the values
  ) 
  
#glimpse(df.bymonth)

p <- ggplot(df.bymonth, aes(x=reviews.month, y=proportion*100, group=car_characteristic, color=car_characteristic)) +
  geom_smooth(method = "loess", span = .25) +
  labs(
  x = "Year", 
  y = "% of Reviews Mentioning Characteristic",
  color = "Car Characteristic"
  ) +
  theme_minimal()

p


```


A few interesting trends are evident. First, driving characteristics and reliability remain the most commonly-referenced characteristics throughout the period, suggesting that consumers highly value these characteristics. Second, comments regarding a car's technology have steadily increased overtime, from a low of 11\% of reviews in 2002 to roughly 35\% of reviews by 2018. Were the data to cover the present period, I would expect this figure to be even higher. Lastly, references to fuel economy closely mirror long-term oil prices:[^note2] references spiked in 2005-06, following a jump in oil prices to above \$80 per gallon. Oil prices (and references to fuel economy) remained elevated through 2014, whereafter we saw a drop to below \$60 per gallon. 

[^note2]: See [**https://www.macrotrends.net/1369/crude-oil-price-history-chart**](https://www.macrotrends.net/1369/crude-oil-price-history-chart)

According to GPT, 70\% of reviews were positive, 24\% were negative, and 6 \% were neutral. Disaggregating results by positive vs. negative review, a number of interesting patterns emerge: 


<!-- 
The dominance of positive reviews is perhaps a little surprising, since behavioral science would predict unsatisfied customers to be the most vocal, along the lines of angry reddit reviews. 
-->


```{r}
#| echo: false
#| warning: false

df.bysentiment.bymonth <- df.coded |>
  mutate(reviews.month = round_date(Review_Date, "month")) |> # reviews by month-year
  filter(sentiment != "neutral") |>
  group_by(reviews.month, sentiment) |>
  summarise(
    #mean.Rating = mean(Rating, na.rm=TRUE),
    #mean.gpt_Rating = mean(gpt_Rating, na.rm=TRUE),
    mean.reliability = mean(reliability, na.rm=TRUE),
    #mean.utility = mean(utility, na.rm=TRUE),
    mean.driving_characteristics = mean(driving, na.rm=TRUE),
    mean.comfort = mean(comfort, na.rm=TRUE),
    #mean.interior = mean(interior, na.rm=TRUE),
    mean.fuel_efficiency = mean(fuel_efficiency, na.rm=TRUE),
    mean.technology = mean(technology, na.rm=TRUE)
  ) |> 
  pivot_longer(
  cols = starts_with("mean."),  # columns to reshape
  names_to = "car_characteristic",          # new column name for the variable names
  names_prefix = "mean.",       # remove prefix from column names
  values_to = "proportion"            # new column name for the values
  ) 
  

ggplot(df.bysentiment.bymonth, aes(x = reviews.month, y = proportion*100, color = car_characteristic)) +
  geom_smooth(method = "loess", span = .25) +
  #geom_point() +
  #facet_grid(~ sentiment) +  # Columns for positive/negative
  facet_grid(
    ~ sentiment, 
    labeller = as_labeller(c("negative" = "Negative Reviews",
                             "positive" = "Positive Reviews"))  # Apply custom labels
  ) +
  labs(
    x = "Year", 
    y = "% of Reviews Mentioning Characteristic",
    color = "Car Characteristic"
  ) +
  theme_minimal()


```

For negative reviews, by far the most frequent complaint relates to reliability. To a lesser degree, unsatisfied customers complain about faulty technology, or unpleasant driving characteristics. Surprisingly, they don't seem to complain too much about poor fuel economy. For positive reviews, satisfied customers most commonly praise the vehicle's driving characteristics, followed by comfort and reliability. References to technology have increased overtime, but remain subordinate through 2018, the latest year for which data are available. 



```{r}
#| eval: false
#| include: false

# reliability
# engine power / performance
# build quality
# smooth handling
# reliability and durability
# technology and/or connectivity
# gas mileage / fuel efficiency
# safety
# cargo space
# driver assistance
# comfort features / car noise / ride quality

# niche issues
# electrical problems
# transmission problems

# other
# resale value
# brand loyalty?

#edmunds categories: utility driving comfort interior technology


```



## Accuracy of LLM-Assisted Coding

We've coded the data and summarized the results, but how confident should we be in the accuracy of GPT's coding? In real-world applications, the best way to verify the accuracy of LLM-coded data would be to compare LLM-generated codings to a randomly-selected test set human-coded reviews. An appropriate sample size for the test set would be in the range of 50 to 100 randomly-sampled, human-coded reviews.[^samplesizenote] But since this is just a demonstration, we won't go through the process human-coding. Instead, as prime-facie validity test, we can take advantage of the fact that reviewers provided a 0-5 satisfaction rating alongside their review; we can compare this rating to GPT's positive-neutral-negative rating, on the assumption that the sentiment of reviewers' reviews matched their satisfaction ratings. For purposes of this exercise, we'll assume ratings below 2.5 had a negative sentiment, ratings between 2.5 and 3.5 had a neutral sentiment, and ratings above 3.5 had a positive sentiment. 


[tbd update to account for the fact that we did indeed prompt gpt to directly coded satisfaction on a scale of 0 to 5.]

[^samplesizenote]: With a sample size of 100, the margin of error for the test error rate for each code/variable would be about 9 percentage points. 


```{r}
#| warning: false


df$GPT_Rating <- df$Rating + rnorm(length(df$Rating),0,1)

# linear trend + confidence interval
p3 <- ggplot(df, aes(x=Rating, y=GPT_Rating)) +
  geom_point() +
  geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
  theme_ipsum()

p3

# confusion matrix approach

# Define the bin breaks and labels
bin_breaks <- c(-Inf, 2.5, 3.5, 5)
bin_labels <- c("Negative", "Neutral", "Positive")

# Use cut() to bin the Rating variable
df <- df %>%
  mutate(Rating_cat = cut(Rating, breaks = bin_breaks, labels = bin_labels, include.lowest = TRUE))



# Load necessary library
library(caret)

# Example data: Actual and predicted classes
actual <- factor(c("Class1", "Class2", "Class3", "Class1", "Class2", "Class3", "Class1", "Class2", "Class3"))
predicted <- factor(c("Class1", "Class2", "Class3", "Class1", "Class2", "Class2", "Class3", "Class3", "Class1"))

# Ensure both actual and predicted have the same levels (classes)
levels(actual) <- levels(predicted) <- c("Class1", "Class2", "Class3")

# Generate the confusion matrix
conf_matrix <- confusionMatrix(predicted, actual)

# Print the confusion matrix
print(conf_matrix)

# Load necessary library
library(ggplot2)

# Convert confusion matrix to a data frame
conf_matrix_df <- as.data.frame(conf_matrix$table)

# Plot the confusion matrix
ggplot(conf_matrix_df, aes(x = Reference, y = Prediction, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), color = "white") +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  labs(title = "Confusion Matrix",
       x = "Actual",
       y = "Predicted") +
  theme_minimal()


```





## Disaggregation

results by American, European, Japanesse, and South Korean automakers





































